SMT-Boosted Security Types for Low-Level MPC

### 2024-10-14 esop25/appendix.tex
checkpoint

\begin{lemma}
  \label{lemma-interference}
  Given $\prog$ with $\iov(\prog) = S \cup V \cup O$ and $\flips(\prog) = R$,
  for all $x \in S \cup R$ and $M \subseteq V$ exactly one of the following conditions holds:
  \begin{enumerate}[\hspace{5mm}i.]
  \item $\notsep{\progd(\prog)}{\setit{x}}{M}$
  \item There exists $X \subseteq S \cup R$ such that
    $\notsep{\progd(\prog)}{\setit{x} \cup X}{M}$ while
    $\sep{\progd(\prog)}{\setit{x}}{M}$ and $\sep{\progd(\prog)}{X}{M}$,
    and we say that $x$ encodes $X$ in $M$.
  \item Neither condition (i) nor (ii) hold, implying $\sep{\progd(\prog)}{\setit{x}}{M}$.
  \end{enumerate}
\end{lemma}

@@

\begin{definition}
  \label{definition-sound}
  Given $\prog$ with $\iov(\prog) = S \cup M \cup P \cup O$
  and $\flips(\prog) = R$, we say that
  \emph{$\Gamma$ is sound for $\prog$} iff for all $M' \subseteq M$
      and $x \in S \cup R$, the following
      conditions hold:
  \begin{enumerate}[\hspace{5mm}i.]
  \item  if $\notsep{\progd(\prog)}{\setit{x}}{M'}$
    there exists $\ty$ with $\leakj{\Gamma}{M'}{\ty}$ and $x \in \ty$.
    \item  if there exists  $X \subseteq S \cup R$ such that
      $\setit{x}$ encodes $X$ in $M'$, 
      then $\leakj{\Gamma}{M'}{\ty}$ and
      $\leakclose{\Gamma}{\ty \cup X}{\ty'}$ and
      $x \in \ty'$.
  \end{enumerate}
\end{definition}

\begin{lemma}
  \label{lemma-cpjsound}
  If $\cpj{R}{\eqs}{\eqspre \wedge \toeq{\prog}}{\Gamma}$ is valid then $\Gamma$ is
  sound for $\prog$.
\end{lemma}
\begin{proof}
  Unrolling Definition \ref{definition-sound}, proceed by induction on $M'$.
  In case $M' = \varnothing$ the result follows trivially. Otherwise
  $M' = M_0 \cup \{ \mx{w}{\cid} \}$ and wlog $\Gamma = \Gamma_0; \mx{w}{\cid} : \ty_0$
  and $\prog = \prog_0;\xassign{\mx{w}{\cid}}{\be}{\cid_0}$, where $R = R_0;R_1$ and:
  \begin{mathpar}
    \cpj{R_0}{\eqs}{\eqspre \wedge \toeq{\prog_0}}{\Gamma_0}

@@

    \eqj{R_2}{\eqs}{R_2}{\ty_1}
  \end{mathpar}
  These facts imply that $\rx{w}{\cid}$ is used as a one-time-pad to encode
  $\vars(\phi)$ in $M'$ 
  \cite{barthe2019probabilistic}. Observing that:
  $$
  \leakclose{\Gamma}{\setit{\cty{\rx{w}{\cid}}{\ty_1}, \rx{w}{\cid}}}{\ty'}
  $$
  The result in this subcase follows by the (second) induction hypothesis. \qed
\end{proof}

\begin{proof}[Theorem \ref{theorem-cpj}]
  Immediate by Lemmas \ref{lemma-cpjsound} and \ref{lemma-interference}. \qed
\end{proof}

@@

\begin{definition}
  Given $H,C$ and $\prog$ with $\iov(\prog) = S \cup V \cup O$,
  we say $x \in V \cup O$ has an \emph{aberration} iff
  $\exists \adversary, \store \in \botruns(\prog)$ such that
  $\store(x) \ne \bot$ and $\neg\exists \store' \in \runs(\prog)$
  with  $\store_{S_H} = \store'_{S_H}$ and
  $\store(x) = \lcod{\store',\be}{\cid}$ where
  $\xassign{x}{\be}{\cid} \in \prog$.
\end{definition}

\begin{lemma}
  Given protocol $\prog$ with
  $\views(\prog) = V$, if 
  $\ipj{\eqs}{\prog}{\Delta}$ is valid
  for some $\eqs$ and $x \in \cinputs$ has an
  aberration for some $H,C$, then $\cheatj{\Delta}{H,C}{\seclev}$
  and $\seclev(x) = \lolab$.
\end{lemma}

\begin{lemma}
  If $\prog$ with $\iov(\prog) = S \cup V \cup O$ does not have
  integrity for some $H,C$, then there exists $\setit{x} \cup V'
  \subseteq \cinputs$ where $\setit{x}$ has an aberration and
  $\notcondsep{\progd(\prog)}{S_H}{\setit{x} \cup V'}{\houtputs \cup O_H}$.
\end{lemma}

\begin{lemma}
  Given $\prog$ with $\iov(\prog) = S \cup V \cup O$ and
  $\ipj{\eqs}{\prog}{\Delta}$ valid for some $\eqs$.  For all $x \in
  \cinputs$ with an aberration, if
  $\notcondsep{\progd(\prog)}{S_H}{\setit{x} \cup V'}{\houtputs \cup O_H}$
  then there exists $y \in \houtputs \cup O_H$ such that
  $\cheatj{\Delta}{H,C}{\seclev}$ and $\seclev(y) = \lolab$.
\end{lemma}

@@

\begin{lemma}
  \label{lemma-fresh}
  Given $w$ and $\cid$ with $\fresh(w,\cid)$, if $\notg{\eqs}[w/y_1,\cid/y_n]$
  is not satisfiable, then there exists no $w',\cid'$ such that
  $\notg{\eqs}[w'/y_1,\cid'/y_n]$ is satisifiable.
\end{lemma}

\begin{proof}
  Let $\eqs = \notg{\eqs}[w/y_1,\cid/y_n]$ and $\eqs' =
  \notg{\eqs}[w'/y_1,\cid'/y_n]$. If $w',\cid'$ are fresh or chosen
  such that $\eqs'$ is an $\alpha$-renaming of $\eqs$, the result is
  immediate.  Otherwise, it must be the case that at least two
  distinct variables $x_1$ and $x_2$ in $\eqs$ are renamed as a single
  variable $x$ in $\eqs'$, given freshness of $w,\cid$. Let
  $\mathit{rename}$ be the function that maps variables in $\eqs$ to
  variables in $\eqs'$ such that $\mathit{rename}(\eqs) = \eqs'$. Let
  $\store'$ be a model of $\eqs'$, and define $\store$ such that:
  $$
  \forall x \in \dom(\vars(E)) . \store(x) \defeq \store'(\mathit{rename}(x)) 
  $$
  Then $\store$ must be a model of $\eqs$, which is a contradiction. \qed
\end{proof}

\begin{lemma}
  \label{lemma-eqs-notg}
  For all $\fresh(\mv_1,\ldots,\mv_n)$, $\notg{E_1}$, $\notg{E_2}$, assuming: 
  \begin{mathpar}
    \notg{\eqs_1}[\mv_1/y_1]\cdots[\mv_n/y_n]  \redx \eqs_1

@@

\begin{proof}
  Given the assumptions, suppose that $E_1' \not\models E_2'$
  is an $\alpha$-renaming of $E_1 \models E_2$. In this
  case the consequence follows immediately. Otherwise,
  suppose on the contrary we can witness
  $E_1' \not\models E_2'$. By definition this means that
  $E_1' \wedge \neg E_2'$ is satisfiable
  but $E_1 \wedge \neg E_2$ is not, which is a contradiction
  by Lemma \ref{lemma-fresh}. \qed 
\end{proof}

\begin{proof}[Theorem \ref{theorem-mtj}]
  (Sketch) The result follows by induction on derivation of the
  main type judgement for $\instr$. The \TirName{Mesg} case follows
  by Lemma \ref{lemma-atj-sound}. The \TirName{Encode} and \TirName{App}
  cases follow by definition of the \TirName{Sig} rule and Lemma
  \ref{lemma-eqs-notg}. 
\end{proof}

>>>>

The following Lemma enumerates critical properties of dependence of
views on input secrets and random variables, and follows by standard
properties of pmfs.
\begin{lemma}
  \label{lemma-interference}
  Given $\prog$ with $\iov(\prog) = S \cup V \cup O$ and $\flips(\prog) = R$,
  for all $x \in S \cup R$ and $M \subseteq V$ exactly one of the following conditions holds:
  \begin{enumerate}[\hspace{5mm}i.]
  \item $\notsep{\progd(\prog)}{\setit{x}}{M}$
  \item There exists $X \subseteq S \cup R$ such that
    $\notsep{\progd(\prog)}{\setit{x} \cup X}{M}$ while
    $\sep{\progd(\prog)}{\setit{x}}{M}$ and $\sep{\progd(\prog)}{X}{M}$,
    and we say that $x$ encodes $X$ in $M$.
  \item Neither condition (i) nor (ii) hold, implying $\sep{\progd(\prog)}{\setit{x}}{M}$.
  \end{enumerate}
\end{lemma}

@@

The following Definition characterizes the crucial representation invariant
preserved by confidentiality type checking. 
\begin{definition}
  \label{definition-sound}
  Given $\prog$ with $\iov(\prog) = S \cup M \cup P \cup O$
  and $\flips(\prog) = R$, we say that
  \emph{$\Gamma$ is sound for $\prog$} iff for all $M' \subseteq M$
      and $x \in S \cup R$, the following
      conditions hold:
  \begin{enumerate}[\hspace{5mm}i.]
  \item  if $\notsep{\progd(\prog)}{\setit{x}}{M'}$
    there exists $\ty$ with $\leakj{\Gamma}{M'}{\ty}$ and $x \in \ty$.
    \item  if there exists  $X \subseteq S \cup R$ such that
      $\setit{x}$ encodes $X$ in $M'$, 
      then $\leakj{\Gamma}{M'}{\ty}$ and
      $\leakclose{\Gamma}{\ty \cup X}{\ty'}$ and
      $x \in \ty'$.
  \end{enumerate}
\end{definition}
Now we can prove that this invariant is preserved by type checking.
\begin{lemma}
  \label{lemma-cpjsound}
  If $\cpj{R}{\eqs}{\eqspre \wedge \toeq{\prog}}{\Gamma}$ is valid then $\Gamma$ is
  sound for $\prog$.
\end{lemma}
\begin{proof}
  Unrolling Definition \ref{definition-sound}, proceed by induction on $M'$.
  In case $M' = \varnothing$ the result follows trivially. Otherwise
  $M' = M_0 \cup \{ \mx{w}{\cid} \}$ and wlog $\Gamma = \Gamma_0; \mx{w}{\cid} : \ty_0$
  and $\prog = \prog_0;\xassign{\mx{w}{\cid}}{\be}{\cid_0}$, where $R = R_0;R_1$ and:
  \begin{mathpar}
    \cpj{R_0}{\eqs}{\eqspre \wedge \toeq{\prog_0}}{\Gamma_0}

@@

    \eqj{R_2}{\eqs}{R_2}{\ty_1}
  \end{mathpar}
  These facts imply that $\rx{w}{\cid}$ is used as a one-time-pad to encode
  $\vars(\phi)$ in $M'$ 
  \cite{barthe2019probabilistic}. Observing that:
  $$
  \leakclose{\Gamma}{\setit{\cty{\rx{w}{\cid}}{\ty_1}, \rx{w}{\cid}}}{\ty'}
  $$
  The result in this subcase follows by the (second) induction hypothesis. \qed
\end{proof}
On the basis of the above, proving the main result is straightforward.
\begin{proof}[Theorem \ref{theorem-cpj}]
  Immediate by Lemmas \ref{lemma-cpjsound} and \ref{lemma-interference}. \qed
\end{proof}

@@

A key observation is that, to undermine integrity adversarial strategies
must be different from passive strategies, in particular at least one
adversarial input must be different than any used in a passive strategy
(an abberation). 
\begin{definition}
  Given $H,C$ and $\prog$ with $\iov(\prog) = S \cup V \cup O$,
  we say $x \in V \cup O$ has an \emph{aberration} iff
  $\exists \adversary, \store \in \botruns(\prog)$ such that
  $\store(x) \ne \bot$ and $\neg\exists \store' \in \runs(\prog)$
  with  $\store_{S_H} = \store'_{S_H}$ and
  $\store(x) = \lcod{\store',\be}{\cid}$ where
  $\xassign{x}{\be}{\cid} \in \prog$.
\end{definition}
It is straightforward to prove that if an abberation can
exist, then it will be assigned $\lolab$ integrity in our
analysis. 
\begin{lemma}
  \label{lemma-aberration-low}
  Given protocol $\prog$ with
  $\views(\prog) = V$, if 
  $\ipj{\eqs}{\prog}{\Delta}$ is valid
  for some $\eqs$ and $x \in \cinputs$ has an
  aberration for some $H,C$, then $\cheatj{\Delta}{H,C}{\seclev}$
  and $\seclev(x) = \lolab$.
\end{lemma}
\begin{proof}
  Given $H,C$, \emph{every} adversarial input is assigned $\lolab$
  integrity by definition of the $\TirName{Mesg}$ rule and the
  \emph{leak} judgement.  The exception to this is messages that pass
  the $\TirName{MAC}$ check-- success of this check guarantees that
  the message is \emph{not} an aberration, so it can safely be
  switched to $\hilab$ integrity. \qed
\end{proof}
The following Lemma follows trivially by Definition \ref{def-integrity},
since without an aberration adversarial inputs are the same as some passive
strategy.
\begin{lemma}
  \label{lemma-undermined-logic}
  If $\prog$ with $\iov(\prog) = S \cup V \cup O$ does not have
  integrity for some $H,C$, then there exists $\setit{x} \cup V'
  \subseteq \cinputs$ where $\setit{x}$ has an aberration and
  $\notcondsep{\progd(\prog)}{S_H}{\setit{x} \cup V'}{\houtputs \cup O_H}$.
\end{lemma}

\begin{lemma}
  \label{lemma-undermined}
  Given $\prog$ with $\iov(\prog) = S \cup V \cup O$ and
  $\ipj{\eqs}{\prog}{\Delta}$ valid for some $\eqs$.  For all $x \in
  \cinputs$ with an aberration, if
  $\notcondsep{\progd(\prog)}{S_H}{\setit{x} \cup V'}{\houtputs \cup O_H}$
  then there exists $y \in \houtputs \cup O_H$ such that
  $\cheatj{\Delta}{H,C}{\seclev}$ and $\seclev(y) = \lolab$.
\end{lemma}
\begin{proof}
  By Lemma \ref{lemma-aberration-low} and definition of the
  confidentiality type analysis, since if $\setit{x} \cup V'$
  interferes with $\houtputs \cup O_H$ then $\setit{x}$
  must occur in the definition of some $y \in \houtputs \cup O_H$,
  yielding $\seclev(y) = \lolab$ in the integrity label assignment
  judgement defined in Figure \ref{fig-cheatj}. \qed
\end{proof}
On the basis of the above we can prove our main result. 
\begin{proof}[Theorem \ref{theorem-ipj}]
  By Lemmas \ref{lemma-undermined} and \ref{lemma-undermined-logic}. \qed
\end{proof}

@@

Soundness of the $\metaprot$ type analysis follows mostly by a
straightforward mapping to the confidentiality and integrity type
systems. The most interesting rules are signature verification
($\TirName{Sig}$) and function application ($\TirName{App}$). 
Here the result hinges on the selection of ``fresh'' variables
that guarantee generalization of constraint entailment.
First we obtain a key result about instantiation and
satisfiability of constraints. 
\begin{lemma}
  \label{lemma-fresh}
  Given $w$ and $\cid$ with $\fresh(w,\cid)$, if $\notg{\eqs}[w/y_1,\cid/y_n]$
  is not satisfiable, then there exists no $w',\cid'$ such that
  $\notg{\eqs}[w'/y_1,\cid'/y_n]$ is satisifiable.
\end{lemma}

\begin{proof}
  Let $\eqs = \notg{\eqs}[w/y_1,\cid/y_n]$ and $\eqs' =
  \notg{\eqs}[w'/y_1,\cid'/y_n]$. If $w',\cid'$ are fresh or chosen
  such that $\eqs'$ is an $\alpha$-renaming of $\eqs$, the result is
  immediate.  Otherwise, it must be the case that at least two
  distinct variables $x_1$ and $x_2$ in $\eqs$ are renamed as a single
  variable $x$ in $\eqs'$, given freshness of $w,\cid$. Let
  $\mathit{rename}$ be the function that maps variables in $\eqs$ to
  variables in $\eqs'$ such that $\mathit{rename}(\eqs) = \eqs'$. Let
  $\store'$ be a model of $\eqs'$, and define $\store$ such that:
  $$
  \forall x \in \dom(\vars(E)) . \store(x) \defeq \store'(\mathit{rename}(x)) 
  $$
  Then $\store$ must be a model of $\eqs$, which is a contradiction. \qed
\end{proof}
Now we can extend the preceding Lemma to constraint entailment,
incorporating evaluation of expressions withing constraints. 
\begin{lemma}
  \label{lemma-eqs-notg}
  For all $\fresh(\mv_1,\ldots,\mv_n)$, $\notg{E_1}$, $\notg{E_2}$, assuming: 
  \begin{mathpar}
    \notg{\eqs_1}[\mv_1/y_1]\cdots[\mv_n/y_n]  \redx \eqs_1

@@

\begin{proof}
  Given the assumptions, suppose that $E_1' \not\models E_2'$
  is an $\alpha$-renaming of $E_1 \models E_2$. In this
  case the consequence follows immediately. Otherwise,
  suppose on the contrary we can witness
  $E_1' \not\models E_2'$. By definition this means that
  $E_1' \wedge \neg E_2'$ is satisfiable
  but $E_1 \wedge \neg E_2$ is not, which is a contradiction
  by Lemma \ref{lemma-fresh}. \qed 
\end{proof}
Now we can consider the main result.
\begin{proof}[Theorem \ref{theorem-mtj}]
  (Sketch) The result follows by induction on derivation of the
  main type judgement for $\instr$. The \TirName{Mesg} case follows
  by Lemma \ref{lemma-atj-sound}. The \TirName{Encode} and \TirName{App}
  cases follow by definition of the \TirName{Sig} rule and Lemma
  \ref{lemma-eqs-notg}. 
\end{proof}
